Run aquaculture model for Atlantic salmon

### 2025-07-22 05_formulating_feeds.qmd
Updates - new feed names

here("src") %>% list.files(pattern = "\\.R$", full.names = TRUE) %>% walk(source)

ingred_inputs_file <- file.path(input_feed_profile_path, "all_ingredients.csv")
feed_input_file <- file.path(input_feed_profile_path, "all_feeds.csv")
```

@@

feed_inputs <- feed_input_file %>% 
  read.csv() %>% 
  pivot_longer(names_to = "feed", values_to = "proportion", cols = !contains(c("ingredient", "proportion"))) %>% 
  mutate(feed = as.factor(feed),
         ingredient = as.factor(ingredient)) %>% 
  merge(ingreds, by = "ingredient", all = T)

>>>>

here("src") %>% list.files(pattern = "\\.R$", full.names = TRUE) %>% purrr::walk(source)

ingred_inputs_file <- file.path(input_feed_profile_path, "all_ingredients.csv")
feed_input_file <- file.path(input_feed_profile_path, "all_feeds.xlsx")
```

@@

feed_inputs <- feed_input_file %>% 
  read_excel(sheet = "testing") %>% 
  pivot_longer(names_to = "feed", values_to = "proportion", cols = !contains(c("ingredient", "category"))) %>% 
  mutate(feed = as.factor(feed),
         ingredient = as.factor(ingredient)) %>% 
  merge(ingreds, by = "ingredient", all = T)

### 2025-07-22 06_run_atlantic_salmon_model.qmd
Updates - new feed names

---
title: "Individual Mariculture Modeling Analysis"
format: html
editor: source
---

@@

This document runs individual mariculture modeling analyses for Atlantic salmon farms. It processes farm location data, species parameters, and conducts sensitivity analyses to understand the impact of various parameters on fish and farm growth measures.

```{r setup}
#| warning: false
#| message: false

library(arrow)
library(sf)
library(dplyr)
library(tidyr)
library(terra)
library(magrittr)
library(furrr)
library(future)
library(tictoc)
library(ggplot2)
library(fs)
library(conflicted)
library(stringr)
library(readxl)
library(units)
library(qs)
library(here)
library(targets)
conflicted::conflicts_prefer(dplyr::filter(), dplyr::select(), .quiet = T)

@@

```{r filenames}
# Set up parallel processing
# plan(multisession, workers = parallelly::availableCores()-1)

@@

```{r farm-coordinates}
times_N <- c("t_start" = 121, "t_end" = 121+547, "dt" = 1)
times_S <- c("t_start" = 274, "t_end" = 274+547, "dt" = 1)

@@

```{r farm-ts-data}
farms_to_omit <- qread(sprintf(file.path(input_farm_coords_path, "%s_farms_to_omit.qs"), this_species))
farm_SST_data <- read_parquet(file.path(input_farm_sst_path, "farm_SST_extracted.parquet"))
farm_IDs <- farm_SST_data %>%
  filter(!farm_id %in% farms_to_omit) %>%
  distinct(farm_id) %>%
  pull(farm_id)

@@

```{r species-parameters}
species_params <- readxl::read_excel(
  path = species_params_excel["file"], 
  sheet = species_params_excel["sheet"]
)
vals <- species_params$Value
names(vals) <- species_params$Quantity
species_params <- vals[!is.na(vals)]
qsave(species_params, species_params_file)
```

@@

```{r population-parameters}
pop_params <- readxl::read_excel(path = pop_params_excel["file"])
vals <- pop_params$Value
names(vals) <- pop_params$Quantity
pop_params <- vals[!is.na(vals)]
qsave(pop_params, pop_params_file)
```

@@

```{r farm-harvest-size}
Sys.setenv(TAR_PROJECT = "project_farmruns")

targets::tar_make(harvest_size)
targets::tar_load(harvest_size)
qsave(harvest_size, farm_harvest_file)

mean(harvest_size$weight)
min(harvest_size$weight)
max(harvest_size$weight)

ggplot(harvest_size, aes(x = weight)) +
  geom_histogram(fill = "salmon", alpha = 0.75, colour = "black")
```

@@

Run the sensitivity runs using the targets pipeline. 
Note that this can take a long time (hours), as the species parameters (19) are run for each farm (2721) and factor (3) and the population parameters (6) are run for a smaller number of farms (271) and factor with a simulated population of 500 fish.

```{r sensitivity-run}
Sys.setenv(TAR_PROJECT = "project_sensitivities")

targets::tar_validate()
targets::tar_visnetwork(targets_only = T)
targets::tar_outdated()
targets::tar_make(
  seconds_meta_append = 120
  )
```

@@

```{r sensitivity-results}
targets::tar_load(sens_results_spec)
targets::tar_load(sens_results_pop)

@@

```{r run-farms-pipeline}
Sys.setenv(TAR_PROJECT = "project_farmruns")

targets::tar_validate()
targets::tar_visnetwork(targets_only = T)
targets::tar_outdated()
targets::tar_make(seconds_meta_append = 90)
```

```{r farm-run-results}
targets::tar_load(cohort_results)
```

>>>>

---
title: "Run aquaculture model for Atlantic salmon"
format: html
editor: source
---

@@

This document runs the aquaculture model for Atlantic salmon farms. It processes farm location data, species parameters, and conducts sensitivity analyses to understand the impact of various parameters on fish and farm growth measures.

```{r}
#| label: setup
#| warning: false
#| message: false

library(arrow)
library(sf)
library(dplyr)
library(tidyr)
library(terra)
library(magrittr)
library(purrr)
library(furrr)
library(future)
library(tictoc)
library(ggplot2)
library(fs)
library(conflicted)
library(stringr)
library(readxl)
library(units)
library(qs)
library(here)
library(targets)
conflicted::conflicts_prefer(dplyr::filter(), dplyr::select(), .quiet = T)

@@

```{r}
#| label: filenames

# Set up parallel processing
# plan(multisession, workers = parallelly::availableCores()-1)

@@

```{r}
#| label: farm-coordinates

times_N <- c("t_start" = 121, "t_end" = 121+547, "dt" = 1)
times_S <- c("t_start" = 274, "t_end" = 274+547, "dt" = 1)

@@

```{r}
#| label: farm-ts-data

farms_to_omit <- qread(sprintf(file.path(input_farm_coords_path, "%s_farms_to_omit.qs"), this_species))
farm_SST_data <- read_parquet(file.path(input_farm_sst_path, "farm_SST_extracted.parquet"))
farm_IDs <- farm_SST_data %>%
  filter(!farm_id %in% farms_to_omit) %>%
  distinct(farm_id) %>%
  pull(farm_id)

@@

```{r}
#| label: species-parameters

species_params <- readxl::read_excel(
  path = species_params_excel["file"], 
  sheet = species_params_excel["sheet"]
)
vals <- species_params$Value
names(vals) <- species_params$Quantity
species_params <- vals[!is.na(vals)]
qsave(species_params, species_params_file)
```

@@

```{r}
#| label: population-parameters

pop_params <- readxl::read_excel(path = pop_params_excel["file"])
vals <- pop_params$Value
names(vals) <- pop_params$Quantity
pop_params <- vals[!is.na(vals)]
qsave(pop_params, pop_params_file)
```

@@

```{r}
#| label: farm-harvest-size

Sys.setenv(TAR_PROJECT = "project_farmruns")

# Check that the feed will actually work
targets::tar_make(test_reference_feed)
targets::tar_load(test_reference_feed)

# Get harvest sizes for all farms
targets::tar_make(harvest_size)
targets::tar_load(harvest_size)

ggplot(harvest_size, aes(x = weight)) +
  geom_histogram(fill = "salmon", alpha = 0.75, colour = "black")

qsave(harvest_size, farm_harvest_file)
```

@@

Run the sensitivity runs using the targets pipeline. 
Note that this can take a long time. 
The species parameters (19) are run for a single fish in each farm (2721) and factor (3), which takes ~30 minutes.
The population parameters (6) are run for a smaller number of farms (271) and each factor (3) with a simulated population of 500 fish, which takes ~30 hours.

```{r}
#| label: sensitivity-run

Sys.setenv(TAR_PROJECT = "project_sensitivities")

targets::tar_validate()
# targets::tar_visnetwork(targets_only = T)
# targets::tar_outdated()
targets::tar_make(
  names = "sens_results_spec",
  seconds_meta_append = 120
  )
```

@@

```{r}
#| label: sensitivity-results

targets::tar_load(sens_results_spec)
targets::tar_load(sens_results_pop)

@@

Note that this also takes a long time - each branch of farmrun takes ~4.5 minutes, and a maximum of ~12 workers can run at a time. The total therefore takes ~17 hours.

```{r}
#| label: run-farms-pipeline

Sys.setenv(TAR_PROJECT = "project_farmruns")

targets::tar_validate()
# targets::tar_visnetwork(targets_only = T)
# targets::tar_outdated()
targets::tar_make(farm_results, seconds_meta_append = 90)
meta <- targets::tar_meta() %>% 
  filter(grepl("farm_run", name))
meta$seconds %>% mean()/60
```

```{r}
#| label: farm-run-results

targets::tar_load(cohort_results)
```
